On the presentation of orthogonal groups over polynomial rings

### 2019-06-08 dl-paper.tex
Added statement of Lemma 8

\subsection{Lemma on the presentation of Steinberg groups}
%Throughout this section $\Phi$ is a root system of type $\rA_{\geq 4}$, $\rD_{\geq 5}$, $\rE_{6,7,8}$ and $A$ is a local ring with maximal ideal $M$.
Let $I\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[X, X\inv]$ as a $\mathbb{Z}$-graded ring in which $X^{-1}$ has degree $1$.
We denote by $B = B(A, I)$ the subring $A[X^{-1}] + I[X] \subseteq A[X, X\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=I \cdot X^d$ for $d>0$, and $B_d=A \cdot X^d$ for $d\leq0$.
Obviously, $B = A[X^{-1}]$ in the case $I=0$ and $B = A[X, X^{-1}]$ in the case $I=A$.

Whenever the coefficient $\xi$ of a Steinberg generator $g = x_\alpha(\xi)$ of $\St(\Phi, B)$ is a homogeneous element of $B$, i.\,e. $\xi \in B_d$ for some $d \in \mathbb{Z}$,
 we call the corresponding generator $g$ homogeneous of degree $d$.
It is not hard to show that $\St(\Phi, B)$ can be presented by the set of all homogeneous Steinberg generators modulo the following set of Steinberg relations
 between them (below $a, a' \in B_d$, $b\in R_e$ and $d,e \in \mathbb{Z}$): 
\begin{align}
&\,\,x_{\alpha}(a)x_{\alpha}(a') =  x_{\alpha}(a+a'),& \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha, \beta} \cdot ab),& \alpha + \beta \in \Phi \tag{R$2_{d,e}$} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha + \beta \not \in \Phi\cup\{0\}.\ \tag{R$3_{d,e}$}
%&[x_{\alpha}(a),\,x_{\beta}(b)]= 1,& \alpha - \beta \in \Phi,\ \tag{R$3^\angle_{d,e}$}
\end{align}
%For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

For $n\geq 1$ we define ``truncated'' Steinberg groups $\St^{\leq n}(\Phi, B)$ by the set of homogeneous Steinberg generators of degree $\leq n$
modulo relations $\text{R}1_{d}$, $\text{R3}_{d, e}$ for all $d,e \leq n$ and $\text{R2}_{d,e}$ for those $d,e \leq n$ for which $d+e \leq n$.

It will be convenient for us to distinguish between the cases when depending on whether the roots $\alpha, \beta$ appearing in $\text{R3}_{d,e}$ are orthogonal to each other or form a sharp angle.
In the former case we denote such relator by $\text{R3}^\bot_{d,e}$, and in the latter case by $\text{R3}^\angle_{d,e}$.

The following lemma is inspired by the proof of the main result in ???.
Basically, it says that most of the relators $\text{R3}^\bot_{d,e}$ in the presentation of $\St^{\leq n}(\Phi, B)$ are superflous.
\begin{lemma}\label{superfluous-relations}

\end{lemma}
\begin{proof}
We will use the following commutator identities (cf Rehmann-Soule).
\begin{align} \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].& \end{align}

@@

In the case $d \leq 0 \leq e \leq n$
\begin{align*} 
 [x_\alpha(\xi), x_{\gamma}(bt^{e})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$ } \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{d,1}$ } \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(b't^{e-1}\xi), x_{-\beta}(t)] & \text{ by $\text{R2}_{d,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$  } \\
 = 1 & \text{ by $\text{R3}^\bot_{d+e-1,1}$. \qedhere} \end{align*}
\end{proof}

@@

Now let us prove $\text{R3}_{1,1}^\angle$. As before, we may assume $p(g, h, u) \in j_+(K(A[X], M[X]) \cap P_\alpha(m) \cap P_{\beta}(m'))$ for some $m, m' \in M$.
Set $\epsilon = N_{\alpha,-\beta}$. Applying~\cref{SR:sharp} we obtain that
\begin{gather*} \begin{multlined} \nonumber
 T_\beta(bX^{-1}) \cdot T_\alpha(aX^{-1}) \left(g, h, u\right) = 
 T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) ,\ u(1 + am)\right) \sim \\ 
 \sim T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am'),\ u(1 + am)\right) = \\
 = (x_\beta(bX^{-1}) x_\alpha(aX^{-1}) \cdot g, h', u(1 + am + bm')), \text{ where} \end{multlined} \\
\begin{multlined} \nonumber
 h' = h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am') \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right) \sim \\
    \sim h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha}\left(\tfrac{m'abX^{-1}(1+am)}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right) = \\
 = h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am} + \tfrac{m'abX^{-1}}{(1+am)(1 + am + bm')}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right) \cdot h^{-1}_\alpha(1 + am) \sim \\
 \sim h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right).
\end{multlined}
\end{gather*}
The expressions in the right-hand sides of the above formulae are clearly symmetric under the action of the transposition $(a, \alpha) \leftrightarrow (b, \beta)$, which implies the desired relation.
\end{proof}

@@

\include{rehmann}
\end{comment}

>>>>

\subsection{Presentation by Homogeneous elements}
%Throughout this section $\Phi$ is a root system of type $\rA_{\geq 4}$, $\rD_{\geq 5}$, $\rE_{6,7,8}$ and $A$ is a local ring with maximal ideal $M$.
Let $I\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[X, X\inv]$ as a $\mathbb{Z}$-graded ring in which $X^{-1}$ has degree $1$.
We denote by $B = B(A, I)$ the subring $A[X^{-1}] + I[X] \subseteq A[X, X\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=I \cdot X^d$ for $d>0$, and $B_d=A \cdot X^d$ for $d\leq0$.
Obviously, $B = A[X^{-1}]$ in the case $I=0$ and $B = A[X, X^{-1}]$ in the case $I=A$.

Whenever the coefficient $\xi$ of a Steinberg generator $g = x_\alpha(\xi)$ of $\St(\Phi, B)$ is a homogeneous element of $B$, i.\,e. $\xi \in B_d$ for some $d \in \mathbb{Z}$,
 we call the corresponding generator $g$ homogeneous of degree $d$.
It is not hard to show that $\St(\Phi, B)$ can be presented by the set of all homogeneous Steinberg generators modulo the following set of Steinberg relations
 between them (below $a, a' \in B_d$, $b\in R_e$ and $d,e \in \mathbb{Z}$): 
\begin{align}
&\,\,x_{\alpha}(a)\cdot x_{\alpha}(a') =  x_{\alpha}(a+a'),                        & \tag{R$1_d$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= x_{\alpha+\beta}(N_{\alpha, \beta} \cdot ab),   & \alpha + \beta \in \Phi,\ \tag{R$2_{d,e}$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= 1,                                              & \alpha - \beta \in \Phi,\ \tag{R$3^\angle_{d,e}$} \\
&\,[x_{\alpha}(a),\,x_{\beta}(b)]= 1,                                              & \alpha \perp \beta \in \Phi.\ \tag{R$3^\bot_{d,e}$}
\end{align}
%For such a ring $R$ one can assume in the definition of the Steinberg group that the generators $x_\alpha(a)$ and relations R1--R3 are defined only for homogeneous elements.

For $n\geq 1$ we define ``truncated'' Steinberg groups $\St^{\leq n}(\Phi, B)$ by the set of homogeneous Steinberg generators of degree $\leq n$
modulo relations $\text{R}1_{d}$, $\text{R3}^\angle_{d, e}$, $\text{R3}_{d,e}^\bot$ for all $d,e \leq n$ and $\text{R2}_{d,e}$ for $d,e,d+e \leq n$.
The following lemma asserts that most of the relators $\text{R3}^\bot_{d,e}$ in this presentation of $\St^{\leq n}(\Phi, B)$ can be omitted.
\begin{lemma}\label{superfluous-relations}
 One can exclude from the presentation of $\St^{\leq n}(\Phi, B)$ all the relators $\text{R3}_{d,e}^\bot$ for which $\max(0,d) + \max(0,e) > 1$.
\end{lemma}
\begin{proof}
We will use the following commutator identities (cf Rehmann-Soule).
\begin{align} \label{eq:H1ii}  [ab, c] = {}^a[b, c] \cdot [a,c];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [a,c]   = 1    \text{ implies } [a, [b,c]] = [[a,b],{}^bc].& \end{align}

@@

In the case $d \leq 0 \leq e \leq n$ for $\xi \in B_d$ one has
\begin{align*} 
 [x_\alpha(\xi), x_{\gamma}(bt^{e})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$ } \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{d,1}$ } \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(b't^{e-1}\xi), x_{-\beta}(t)] & \text{ by $\text{R2}_{d,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$  } \\
 = 1 & \text{ by $\text{R3}^\bot_{d+e-1,1}$. \qedhere} \end{align*}
\end{proof}

@@

Now let us prove $\text{R3}_{1,1}^\angle$. As before, we may assume $p(g, h, u) \in j_+(K(A[X], M[X]) \cap P_\alpha(m) \cap P_{\beta}(m'))$ for some $m, m' \in M$.
Set $\epsilon = N_{\alpha,-\beta}$. Applying~\cref{SR:sharp} we obtain that
\begin{gather*} \begin{multlined} \nonumber
 T_\beta(bX^{-1}) \cdot T_\alpha(aX^{-1}) \left(g, h, u\right) = 
 T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) ,\ u(1 + am)\right) \sim \\ 
 \sim T_\beta(bX^{-1}) \left(x_\alpha(aX^{-1})\cdot g,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am'),\ u(1 + am)\right) = \\
 = (x_\beta(bX^{-1}) x_\alpha(aX^{-1}) \cdot g, h', u(1 + am + bm')), \text{ where} \end{multlined} \\
\begin{multlined} \nonumber
 h' = h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-\epsilon am') \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right) \sim \\
    \sim h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right) \cdot h^{-1}_\alpha(1 + am) \cdot x_{\alpha}\left(\tfrac{m'abX^{-1}(1+am)}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}(1+am)}{1 + am + bm'}\right) = \\
 = h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am} + \tfrac{m'abX^{-1}}{(1+am)(1 + am + bm')}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right) \cdot h^{-1}_\alpha(1 + am) \sim \\
 \sim h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am + bm'}\right) \cdot x_\beta\left(-\tfrac{bX^{-1}}{1 + am + bm'}\right).
\end{multlined}
\end{gather*}
The expressions in the right-hand sides of the above formulae are clearly symmetric under the action of the transposition $(a, \alpha) \leftrightarrow (b, \beta)$, which shows that the desired relation holds.
\end{proof}

@@

\end{comment}
\include{rehmann}
